material and methods

### 2018-03-26 man/ms_1_material_methods.Rmd
correct passive and active voice methods

## Greenness data 
As a proxy of primary growth, we characterized the vegetation greenness of *Q. pyrenaica* by means of the *Enhanced Vegetation Index* (EVI) derived from MOD13Q1 product obtained by the *Moderate Resolution Imaging Spectroradiometer* (MODIS) sensor [@Didan2015]. Vegetation indices derived from remote sensing information are linked with primary productivity [@Ruimy1994] and they have been been widely used to derive descriptors of ecosystem functioning [*e.g* @AlcarazSegura2009]. These indices are linearly related to net primary productivity through the fraction of photosynthetically active radiation intercepted by green vegetation (*i.e.* vegetation greenness) [@Ruimy1994].

We used EVI instead of NDVI (*Normalized Difference Vegetation Index*; another of the most common greenness vegetation indices) because EVI is more sensitive to changes in high-biomass areas (a serious shortcoming of NDVI); EVI reduces the influence of atmospheric conditions on vegetation index values, and EVI corrects for canopy background signals [@Huete2002; @Krapivin2015; @Cabello2012].

MODIS EVI Data from Collection 6 were obtanied using Google Earth Engine platform for the period 2000 - 2016. EVI data consits of 16-day maximun value composite images (23 per year) of the EVI value with a spatial resolution of 250 m x 250 m. We selected the pixels covering the distribution of *Q. pyrenaica* forests in Sierra Nevada (*n* = 928 pixels). A data filtering was applied to select EVI valid values. The filtering was done using quality flags and VI Usefulness Indices accompanying the EVI data. We filter out those values affected by high content of aerosols, clouds, snow and shadows, following the filtering recommendations of EVI data for mountain regions [@Reyes2015].

>>>>

## Greenness data

Vegetation indices derived from remote sensing information have been widely used to derive descriptors of ecosystem functioning [*e.g* @AlcarazSegura2009] and they are linked with primary productivity [@Ruimy1994]. Here, vegetation greenness of *Q. pyrenaica* was characterized by means of the *Enhanced Vegetation Index* (EVI), derived from MOD13Q1 product of the *Moderate Resolution Imaging Spectroradiometer* (MODIS) sensor [@Didan2015]. MODIS EVI Data (Collection 6) were obtanied using Google Earth Engine platform for the period 2000 - 2016. EVI data consits of 16-day maximun value composite images (23 per year) of the EVI value with a spatial resolution of 250 m x 250 m. We selected the pixels covering the distribution of *Q. pyrenaica* forests in Sierra Nevada (*n* = 928 pixels). Then, a data filtering was applied to select EVI valid values. For this purpose, complementary information accompanying the EVI data was used (VI Usefulness Indices and quality flags). Values affected by high content of aerosols, clouds, snow and shadows, were filtered out following recommendations of @Reyes2015 for mountain regions [@Reyes2015].

```{r echo=FALSE, results='asis', eval=show_text}
cat("
We used EVI instead of NDVI (*Normalized Difference Vegetation Index*; another of the most common greenness vegetation indices) because EVI is more sensitive to changes in high-biomass areas (a serious shortcoming of NDVI); EVI reduces the influence of atmospheric conditions on vegetation index values, and EVI corrects for canopy background signals [@Huete2002; @Krapivin2015; @Cabello2012].
